---
title: "Practical Machine Learning Assignment"
author: "DTSK"
date: "1/19/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Introduction

One thing that people regularly do is quantify how  much of a particular activity they do, but they rarely quantify how well they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants.The goal of this project is to predict the manner in which they did the exercise. 

Preparation

```{r load-packages, echo=TRUE}
library(caret)
library(ggplot2)
library(lattice)
library (rpart)
library(rpart.plot)
library(randomForest)
library(RColorBrewer)
library(corrplot)

```

Downloading of data
```{r, echo=TRUE}
UrlTrain <- "http://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
UrlTest  <- "http://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"

# download the datasets
training <- read.csv(url(UrlTrain))
testing  <- read.csv(url(UrlTest))

# create a partition with the training dataset 
inTrain  <- createDataPartition(training$classe, p=0.7, list=FALSE)
TrainSet <- training[inTrain, ]
TestSet  <- training[-inTrain, ]
dim(TrainSet)
```
```{r, echo=TRUE}
dim(TestSet)
```

Data Preparation

Elimate Variables which are having nearly zero variance.
```{r, echo=TRUE}
NZV <- nearZeroVar(TrainSet)
TrainSet <- TrainSet[, -NZV]
TestSet  <- TestSet[, -NZV]
dim(TrainSet)
```
```{r, echo=TRUE}
dim(TestSet)
```
Remove variable that are NA

```{r, echo=TRUE}
AllNA    <- sapply(TrainSet, function(x) mean(is.na(x))) > 0.95
TrainSet <- TrainSet[, AllNA==FALSE]
TestSet  <- TestSet[, AllNA==FALSE]
dim(TrainSet)
```
```{r, echo=TRUE}
dim(TestSet)
```

Exclude col1 to col5 as they are not related to the model
```{r, echo=TRUE}
TrainSet <- TrainSet[, -(1:5)]
TestSet  <- TestSet[, -(1:5)]
dim(TrainSet)
```
```{r, echo=TRUE}
dim(TestSet)
```
Random Forest
```{r, echo=TRUE}
set.seed(111)
controlRF <- trainControl(method="cv", number=3, verboseIter=FALSE)
modFitRandForest <- train(classe ~ ., data=TrainSet, method="rf",
                          trControl=controlRF)
modFitRandForest$finalModel
```

Prediction on test data
```{r, echo=TRUE}
predictRandForest <- predict(modFitRandForest, newdata=TestSet)
confMatRandForest <- confusionMatrix(predictRandForest, as.factor(TestSet$classe))
confMatRandForest
```


plot matrix results
```{r, echo=TRUE}
plot(confMatRandForest$table, col = confMatRandForest$byClass, 
     main = paste("Random Forest - Accuracy =",
                  round(confMatRandForest$overall['Accuracy'], 4)))
```
Generalized Boosted Model
```{r, echo=TRUE}
set.seed(111)
library(gbm)
```
```{r, echo=TRUE}
controlGBM <- trainControl(method = "repeatedcv", number = 5, repeats = 1)
modFitGBM  <- train(classe ~ ., data=TrainSet, method = "gbm",
                    trControl = controlGBM, verbose = FALSE)
modFitGBM$finalModel
```

Prediction on test dataset
```{r, echo=TRUE}
predictGBM <- predict(modFitGBM, newdata=TestSet)
confMatGBM <- confusionMatrix(predictGBM, as.factor(TestSet$classe))
confMatGBM
```
plot matrix results
```{r, echo=TRUE}
plot(confMatGBM$table, col = confMatGBM$byClass, 
     main = paste("GBM - Accuracy =", round(confMatGBM$overall['Accuracy'], 4)))
```
Applying the Selected Model to the Test Data
The accuracy of the 2 regression modeling methods above are: Random Forest : 0.9978 GBM : 0.9884 In that case, the Random Forest model will be applied to predict the quiz.
```{r, echo=TRUE}
predictTEST <- predict(modFitRandForest, newdata=testing)
predictTEST
```
